---
title: "BBWA (Bay-breasted Warbler)"
format:
  html:
    self-contained: true
    link-external-icon: true
    link-external-newwindow: true
toc: true
date-modified: last-modified
editor: visual
css: styles.css
---

Please add your feedback on each section of the model cycle.

Use this spreadsheet to add your comments..........

```{r echo=F, message=F, warning=FALSE}
# Setting working directory
setwd("C:/FiresideChat")
# Libraries
library(dplyr)
library(reshape2)
library(tidyr)
library(kableExtra)
library(sf)
library(terra)
library(tidyterra)
library(tibble)
library(ggplot2)
library(leaflet)
library(DT)
library(gbm)
library(wrapr)
library(reactable)



```

```{r echo=FALSE, eval=FALSE}

# Setting working directory
setwd("C:/FiresideChat/")

#Loading master file and selecting ecoregions c71 & c81
load("./data/04_NM5.0_data_stratify.R")

#Transofrm bird objec from matrix to df
bird_d <- as.data.frame(as.matrix(bird))

# Saving single .cvs files from the R.Dataset (too big and reduce PC performance)
write.csv(bcrlist,"./data/bcrlist.csv" )
write.csv(covlist, "./data/covlist.csv")
write.csv(bird_d, "./data.bird_d.csv")



```

::: panel-tabset
## 1. Species

### Histogram of counts

This is an histogram of raw count from data sets included in the model.

::: callout-important
## Task:

We would like to know if there are outliers and/or a truncation value that modelers might be aware of.
:::

```{r echo=FALSE, eval=F}

#loading csv files (these are small)
bcrlist <- read.csv("./data/bcrlist.csv")
bird_d <- read.csv("./data/bird_d.csv")

# Selecting BCRs 71 and 81 that are TRUE
bcrlist_sel <- dplyr::select(bcrlist, id, can71, can81)%>%
dplyr::filter(can71 == "TRUE" & can81=="TRUE")

# Merging bcrlist_sel and bird_d by id

 
```

```{r echo=FALSE, eval=FALSE}



#merge_birds <- merge(bcrlist_sel, bird_d)

```

### Temporal distribution of detections (time relative to sunrise, ordinal day)

This is the temporal distribution of detections.

::: callout-important
## Task: 

We would like to know if there are outliers and/or a truncation value that modelers might be aware of.
:::

### Training data

This is the training data.........

::: callout-important
## Task:

Please identify misidentifications, migratory individuals, or surprises
:::

## 2. Covariates

```{r echo=FALSE}
covlist <- read.csv("./data/covlist.csv")
#covlist_sel <- read.csv("./covlist_sel.csv")


covlist_sel <- dplyr::filter(covlist, bcr %in% c("can71", "can81"))%>%
  dplyr::select(where(~ last(.x) != "FALSE"))%>%
  reshape2::melt(id.vars =  "bcr")
 names(covlist_sel)[2] <- "Covariate.label"

cov_label <- read.csv("./data/covariates_label_insert.csv", header=T)

merge_cov <- merge(covlist_sel, cov_label, by="Covariate.label")
cov_bcr71 <- filter(merge_cov, bcr == "can71")
cov_bcr81 <- filter(merge_cov, bcr == "can81")


```

### Covariates relative importance BCR 71 (mean 10 bootstraps)

::: callout-important
## Task

Please re-rank covariates and/or suggest expected relative importance and/or percentage (%).
:::

```{r echo=FALSE, warning=FALSE, message=FALSE}

list_bootstraps <- list.files("./data/BBWA/covariates_bootstrps/can_71",
                             pattern=".R",
                             full.names = T)

df_bootstrap <- do.call(rbind,
               lapply(list_bootstraps, function(x) {
                 
                 load(file = x)
                 gbm::summary.gbm(b.i, plotit = F)
               }))

df_bootstrap_mean <- dplyr::summarise(df_bootstrap, .by=var, mean=mean(rel.inf))



```

```{r echo=FALSE, fig.height=6}

ggplot(data=df_bootstrap_mean, aes(x=reorder(var, mean), y=mean, fill=mean))+
           geom_bar(stat = "identity", fill="gray") + 
   labs(title = "Covariates relative importance",
        x = "Covariates", y = "%")+
  coord_flip()+
    theme(legend.position = "none")
```

#### Covariates list an description

::: callout-important
## Task:

Task: Please suggest adding/removing covariates from the list.
:::

```{r echo=FALSE}

 kable(cov_bcr71, format="html", escape = F, align = "c")%>%
     kableExtra::kable_paper("hover", full_width = F) %>%
  kableExtra::scroll_box(width = "100%", height = "450px")%>%
  kableExtra::kable_styling(position = "center")
```

### Covariates relative importance BCR 81 (mean 10 boostraps)

::: callout-important
## Task:

Please re-rank covariates and/or suggest expected relative importance and/or percentage (%).
:::

```{r echo=FALSE}

list_bootstraps <- list.files("./data/BBWA/covariates_bootstrps/can_81",
                             pattern=".R",
                             full.names = T)

df_bootstrap <- do.call(rbind,
               lapply(list_bootstraps, function(x) {
                 
                 load(file = x)
                gbm::summary.gbm(b.i, plotit = F)
               }))

df_bootstrap_mean <- dplyr::summarise(df_bootstrap, .by=var, mean=mean(rel.inf))



```

```{r echo=FALSE, fig.height=6}

ggplot(data=df_bootstrap_mean, aes(x=reorder(var, mean), y=mean, fill=mean))+
           geom_bar(stat = "identity", fill="gray") + 
   labs(title = "Covariates relative importance",
        x = "Covariates", y = "%")+
  coord_flip()+
    theme(legend.position = "none")
```

#### Covariates list an description

::: callout-important
## Task:

Please suggest adding/removing covariates from the list.
:::

```{r echo=FALSE}

 kable(cov_bcr81, format="html", escape = F, align = "c")%>%
     kableExtra::kable_paper("hover", full_width = F) %>%
  kableExtra::scroll_box(width = "100%", height = "450px")%>%
  kableExtra::kable_styling(position = "center")
```

```{r, echo=FALSE}
reactable <- function(...) {
  htmltools::tagList(reactable::reactable(...))
  
}
```

```{r echo=FALSE}
cov_table <- reactable(cov_bcr81,  highlight = TRUE, defaultPageSize = 4, searchable = TRUE,
          defaultSorted = c("Covariate.label", "Category", "Description"),
          filterable = TRUE)
```

```{r, warning=FALSE, message=FALSE}
cov_table
```

## 3. Data Preparation

## 4. Model fitting

## 5. Predictions

#### Mean density map

This map shows estimates of the number of males per hectare.

::: callout-important
## Task:

Please identify areas that are too high or too low density, if any.
:::

```{r echo=FALSE}

bcr_boundary <- terra::vect("./data/BAM_BCR_NationalModel.shp")

BBWA_can71_2020 <- terra::rast("./data/BBWA/BBWA_can71_2020.tiff")

BBWA_can81_2020 <- terra::rast("./data/BBWA/BBWA_can81_2020.tiff")



```

```{r echo=FALSE, warning=FALSE, message=FALSE}

ggplot(bcr_boundary[bcr_boundary$subUnit==71])+
  geom_spatraster(data=BBWA_can71_2020$mean)+
   geom_spatvector(fill = NA)+
   theme_minimal() +
   scale_fill_whitebox_c(
    palette = "deep",
    n.breaks = 4
  )+
   labs(
    fill = "Mean",
    title = "Spatial predictions",
    subtitle = "BBWA-Bay-breasted Warbler, BCR-71"
  )
```

```{r echo=FALSE, warning=FALSE, message=FALSE}

ggplot(bcr_boundary[bcr_boundary$subUnit==81])+
  geom_spatraster(data=BBWA_can81_2020$mean)+
   geom_spatvector(fill = NA)+
   theme_minimal() +
   scale_fill_whitebox_c(
    palette = "deep",
    n.breaks = 4
  )+
   labs(
    fill = "Mean",
    title = "Spatial uncertainty",
    subtitle = "BBWA-Bay-breasted Warbler, BCR-81"
  )
```

#### Extrapolation map

This map shows.................. \#####

::: callout-important
## Task:

Please identify any areas that are surprising for you.
:::

```{r echo=FALSE, warning=FALSE, message=FALSE}

ggplot(bcr_boundary[bcr_boundary$subUnit==71])+
 geom_spatraster(data=BBWA_can71_2020$extrapolation)+
   geom_spatvector(fill = NA)+
   theme_minimal() +
   scale_fill_whitebox_c(
    palette = "deep",
    n.breaks = 2
  )+
   labs(
    fill = "Binary",
    title = "Extrapolation map",
    subtitle = "BBWA-Bay-breasted Warbler, BCR-71"
  )
```

```{r echo=FALSE, warning=FALSE, message=FALSE}

ggplot(bcr_boundary[bcr_boundary$subUnit==81])+
 geom_spatraster(data=BBWA_can81_2020$extrapolation)+
   geom_spatvector(fill = NA)+
   theme_minimal() +
   scale_fill_whitebox_c(
    palette = "deep",
    n.breaks = 2
  )+
   labs(
    fill = "Binary",
    title = "Extrapolation map",
    subtitle = "BBWA-Bay-breasted Warbler, BCR-81"
  )
```

## 6. Uncertainty

#### Coefficient of variation

::: callout-important
## Task;

Please identify any areas that are surprising for you.
:::

```{r echo=FALSE, warning=FALSE, message=FALSE}

ggplot(bcr_boundary[bcr_boundary$subUnit==71])+
  geom_spatraster(data=BBWA_can71_2020$cv)+
   geom_spatvector(fill = NA)+
   theme_minimal() +
   scale_fill_whitebox_c(
    palette = "muted",
    n.breaks = 5
  )+
   labs(
    fill = "Coefficient \nof variation",
    title = "Spatial uncertainty",
    subtitle = "BBWA-Bay-breasted Warbler, BCR-71"
  )
```

```{r echo=FALSE, warning=FALSE, message=FALSE}

ggplot(bcr_boundary[bcr_boundary$subUnit==81])+
  geom_spatraster(data=BBWA_can81_2020$cv)+
   geom_spatvector(fill = NA)+
   theme_minimal() +
   scale_fill_whitebox_c(
    palette = "muted",
    n.breaks = 5
  )+
   labs(
    fill = "Coefficient \nof variation",
    title = "Spatial uncertainty",
    subtitle = "BBWA-Bay-breasted Warbler, BCR-81"
  )
```
:::
