---
title: "BBWA (Bay-breasted Warbler): Region BCR-71"
format:
  html:
    self-contained: true
    link-external-icon: true
    link-external-newwindow: true
    hide:
      - toc
      - navigation
editor: visual
css: styles.css
---

Welcome to bird model evaluation! Below are the model product inputs and outputs for evaluation, spread across six numbered categories. Within, each category there will be several numbered tasks. Note that some categories are in-progress. Please use this google form [BBWA-BCR-71](https://docs.google.com/forms/d/1xhekH7fj2461LX_AAbN8Md1mLibKmaGo4mF_FexQm3M/edit) to provide your feedback on each section of the model cycle.

```{r echo=F, message=F, warning=FALSE}
# Setting working directory
#setwd("C:/FiresideChat")
# Libraries
library(dplyr)
library(reshape2)
library(tidyr)
library(kableExtra)
library(sf)
library(terra)
library(tidyterra)
library(tibble)
library(ggplot2)
library(leaflet)
library(DT)
library(gbm)
library(wrapr)
library(reactable)
library(lubridate)
library(openxlsx)
library(shiny)
library(widgetframe)
library(leafem)
library(tidyverse)

# Setting working directory
setwd("C:/FiresideChat/")

```

```{r echo=FALSE, warning=FALSE, message=FALSE}
# TASK 1. HISTOGRAM COUNTS
#Loading master file and selecting ecoregions c71 & c81
load("./data/04_NM5.0_data_stratify_7181.R")

# Selecting BCRs 71 and 81 that are TRUE
bcrlist_can71 <- dplyr::select(bcrlist, id, can71)%>%
dplyr::filter(can71 == "TRUE")

bcrlist_can81 <- dplyr::select(bcrlist, id, can81)%>%
dplyr::filter(can81 == "TRUE")

#Transofrm bird objec from matrix to df
bird_d <- as.data.frame(as.matrix(bird))
names(bird_d)[1] <- "id"

# Merging bcrlist_sel and bird_d by id, indicate if species was detected or not
birds_merged_can71 <- merge(bcrlist_can71, bird_d, by="id")%>%
   subset(select = c("id",  "BBWA")) |> 
    mutate(BBWA=ifelse(BBWA=="0", NA, as.integer(BBWA)),
           detection = ifelse(is.na(BBWA), 0, 1)) |> 
  rename(count = BBWA)
  
birds_merged_can81 <- merge(bcrlist_can81, bird_d, by="id")%>%
   subset(select = c("id", "BBWA")) |> 
    mutate(BBWA=ifelse(BBWA=="0", NA, as.integer(BBWA)),
           detection = ifelse(is.na(BBWA), 0, 1)) |> 
  rename(count = BBWA)

# TASK 2. TEMPORAL DISTRIBUTION OF DETECTIONS
# adding the DATE, TIME, and COORDS
visit_detections_can71 <- merge(bcrlist_can71, visit, by="id")%>%
  inner_join(birds_merged_can71) |> 
   subset(select = c("id", "tssr", "jday", "count", "detection", "lat", "lon")) |> 
  pivot_longer(c(jday, tssr), names_to="metric", values_to="value") |> 
  mutate(metric = factor(metric, levels = c("tssr", "jday"),
                            labels = c("Time since sunrise", "Ordinal day")),
         detection = factor(detection, levels=c(0, 1), labels = c("Undetected", "Detected")))

visit_detections_can81 <- merge(bcrlist_can81, visit, by="id")%>%
    inner_join(birds_merged_can71) |> 
   subset(select = c("id", "tssr", "jday", "count", "detection", "lat", "lon")) |> 
  pivot_longer(c(jday, tssr), names_to="metric", values_to="value") |> 
  mutate(metric = factor(metric, levels = c("tssr", "jday"),
                            labels = c("Time since sunrise", "Ordinal day")),
         detection = factor(detection, levels=c(0, 1), labels = c("Undetected", "Detected")))

```

```{r echo=FALSE}

# Boundaries regions
bcr_boundary <- terra::vect("./data/BAM_BCR_NationalModel.shp")

# Boundary geographic for interactive leaflet
bcr_boundary_wgs84 <- terra::project(bcr_boundary, "epsg:4326")

# raster stack
BBWA_can71_2020 <- terra::rast("./data/BBWA/BBWA_can71_2020.tiff")
BBWA_can81_2020 <- terra::rast("./data/BBWA/BBWA_can81_2020.tiff")


```

-   ::: panel-tabset
    # 1. MODEL RATIONALE

    The BAM V5 models were produced for the purpose of modelling the distribution and abundance of breeding birds in the boreal forest. The primary purposes of the models are to accurately map the distribution of bird density on the landscape and sum those predictions to estimate population size.

    ## 2. SPECIES INPUT DATA

    ::: {.callout-note collapse="true"}
    ## a. Histogram of counts

    ::: {style="display: flex;"}
    <div>

    **Task:**

    Please identify any potential outliers and/or a truncation value that modelers should be aware of.

    </div>

    <div>

    **Description:**

    This is a histogram of raw counts included in the model.

    ```{r echo=FALSE, message=FALSE, warning=FALSE}

    # Histogram

    ggplot(dplyr::filter(birds_merged_can71, count > 0), aes(x=count))+
      geom_histogram()+
      labs(x="Count of birds observed", y = "Number of surveys with detections")+
      theme(axis.text=element_text(size=14))+
      theme(axis.title = element_text(size = 16))
     
    ```

    </div>
    :::
    :::

    ::: {.callout-note collapse="true"}
    ## b. Temporal distribution of detections

    ::: {style="display: flex;"}
    <div>

    **Task:**

    Please identify any potential outliers and/or a truncation value that modelers should be aware of.

    </div>

    <div>

    **Description:**

    This is the temporal distribution of detections and nondetections by time since sunrise and ordinal day of year.

    ```{r echo=FALSE, message=FALSE, warning=FALSE}
    # Histogram

    ggplot(visit_detections_can71, aes(x=value))+
      geom_histogram()+
      labs(x="", y = "Number of surveys")+
      facet_grid(detection~metric, scales="free") +
      theme(axis.text=element_text(size=14))+
      theme(axis.title = element_text(size = 16))


     
    ```

    </div>
    :::
    :::

    ::: {.callout-note collapse="true"}
    ## c. Map of observations

    ::: {style="display: flex;"}
    <div>

    **Task:**

    Select "Observations" on the map below. Please identify any observations that might be misidentifications, migratory individuals, or surprises.

    </div>

    <div>

    **Description:**

    Please use the map below to visualize observations of the focal species.

    ```{r echo=FALSE}

    locations_can71 <- st_as_sf(visit_detections_can71, coords=c("lon", "lat"), crs=4326) |> 
      terra::vect() |> 
      terra::project(crs(bcr_boundary))

    locations_can81 <- st_as_sf(visit_detections_can81, coords=c("lon", "lat"), crs=4326) |> 
      terra::vect() |> 
      terra::project(crs(bcr_boundary))


    ```

    </div>
    :::
    :::

    ## 3. COVARIATES

    ```{r echo=FALSE, message=FALSE, warning=FALSE}
    #covlist <- read.csv("./data/covlist.csv")
    #covlist_sel <- read.csv("./covlist_sel.csv")

    covlist_sel <- dplyr::filter(covlist, bcr %in% c("can71", "can81"))%>%
      dplyr::select(where(~ last(.x) != "FALSE"))%>%
      reshape2::melt(id.vars =  "bcr")
     names(covlist_sel)[2] <- "Covariate.label"

    cov_label <- read.xlsx("./data/covariates_label_insert.xlsx")

    merge_cov <- merge(covlist_sel, cov_label, by="Covariate.label")
    cov_bcr71 <- filter(merge_cov, bcr == "can71")%>%
      dplyr::select(Covariate.label, Category, Description)
    cov_bcr81 <- filter(merge_cov, bcr == "can81")

    ```

    ::: {.callout-note collapse="true"}
    ## a. Covariate relative importance

    **Task:** Please re-rank covariates and/or suggest expected % relative importance.

    **Description:** Mean relative importance across 10 bootstraps; error bars represent standard error. Definitions for covariate names can be found in task 2a above.

    ```{r echo=FALSE, warning=FALSE, message=FALSE}

    list_bootstraps <- list.files("./data/BBWA/covariates_bootstrps/can_71",
                                 pattern=".R",
                                 full.names = T)

    df_bootstrap <- do.call(rbind,
                   lapply(list_bootstraps, function(x) {
                     
                     load(file = x)
                     gbm::summary.gbm(b.i, plotit = F)
                   }))

    df_bootstrap_mean <- dplyr::summarise(df_bootstrap, .by=var, mean=mean(rel.inf))

    df_bootstrap_sd <- dplyr::summarise(df_bootstrap, .by=var, sd=sd(rel.inf))

    n = length(df_bootstrap$rel.inf)

    df_bootstrap_sd$se <- df_bootstrap_sd$sd / sqrt(n)

    df_bootstrap_merge <- merge(df_bootstrap_mean, df_bootstrap_sd, by="var")

    ```

    ```{r echo=FALSE, fig.height=6}

    ggplot(data=df_bootstrap_merge, aes(x=reorder(var, mean), y=mean, fill=mean))+
               geom_bar(stat = "identity", fill="gray") + 
       labs(title = "Covariates relative importance",
            x = "Covariates", y = "%")+
      coord_flip()+
        theme(legend.position = "none")+
      geom_errorbar(
        aes(ymin = mean - se, ymax = mean + se), 
        position = position_dodge2(padding = 0.5))
    ```
    :::

    ::: {.callout-note collapse="true"}
    ## b. Covariates list and description

    **Task:** Please suggest adding/removing covariates from the list.

    **Description:** Metadata and preparation for covariates used in model fitting.

    ```{r echo=FALSE}


    cov_table <- htmltools::tagList(DT::datatable(cov_bcr71, filter='top',  rownames = F, escape = F), options = list(
      pageLength = 5, autoWidth = TRUE
    ))
    ```
    :::

    ```{r}
    cov_table
    ```

    ## 4. DATA PREPARATION

    ## a. Sampling data

    **Task:**

    Please identify any sampling data preparation steps that might violate assumptions and suggest alternatives.

    **Description:**

    The following treatments were applied to the survey data

    -   Survey data were spatially thinned using a hexagonal grid with cells of 2.5 km spacing to ensure independence of samples. One survey per grid per year was used for each separate model run.

    -   Survey data were assigned to bird conservation region (BCR) for each modelling unit using all surveys within the BCR and a 100 km buffer around the BCR.

    ## b. Covariate data

    **Task:**

    Please identify any covariate data preparation steps that might violate assumptions and suggest alternatives.

    **Description:**

    The following preparations were made to the covariate data:

    -   Local covariates were extracted with a 200m buffer radius. Landscape covariates were extracted with a 2 km buffer radius.

    -   Covariates were thinned for each BCR using stepwise variance inflation factor (VIF) with a threshold of 10 to reduce covariance, which typically results in no two-way correlation greater than 0.7.

    -   Additional preparations are listed in the table in task 2a.

    ## 5. MODEL FITTING

    ## a. Model process

    **Task:**

    Please identify any modelling choices that might violate assumptions and suggest alternatives.

    **Description:**

    Separate models were fit for each combination of species and BCR. All models were fit with boosted regression trees that could accommodate up to 3-way interactions between variables. The modelling process consisted with the following steps:

    1.  Tuning: models were fit with decreasing learning rates until a reasonable number of trees was found (between 1,000 and 10,000). Covariates with less than 0.1% variable importance were removed.

    2.  Bootstrapping: ten models with different selections of spatial thinning were fit using the learning rate and number of trees defined in step 1.

    3.  Predicting: predictions were made using covariate values for every 5 years from 1985 to 2020 from the bootstrapped models to 1km resolution.

    4.  Validating: values were extracted for withheld data from the prediction surfaces of the closest year and model performance was evaluated for various parameters

    5.  Range limiting: predictions outside of probable range based on eBird detections were masked out.

    ## 6. PREDICTIONS

    ::: {.callout-note collapse="true"}
    ## a. Mean density map

    **Task:**

    Please identify areas that are too high or too low density, if any.

    **Description:**

    Select "Prediction" on the map below. It shows the mean estimates of the number of male birds per hectare across 10 bootstraps.
    :::

    ## 7. UNCERTAINTY

    ::: {.callout-note collapse="true"}
    ## a. Coefficient of variation

    **Task:**

    Please identify areas that have surprisingly high or low uncertainty, if any.

    **Description:**

    Select "Uncertainty" on the map below. It shows the coefficient of variation of the number of male birds per hectare across 10 bootstraps.
    :::

    ::: {.callout-note collapse="true"}
    ## b. Extrapolation map

    **Task:**

    Please identify any areas that are surprising for you.

    **Description:**

    Extrapolation represents ......and was calculated ....

    This map below shows the (EXTRAPOLTAION layer)..................
    :::
    :::

## MAPS

```{r echo=FALSE, warning=FALSE, message=FALSE}

#  SPECIES DATA

  # Pallete
    pal <- colorNumeric(palette = "viridis",domain = locations_can71$count)
    pal2 <- colorQuantile("Greens", values(BBWA_can71_2020$mean),
                          na.color = "transparent")
    pal3 <- colorQuantile("Blues", values(BBWA_can71_2020$cv),
                          na.color = "transparent")
    pal4 <- colorNumeric("Reds", values(BBWA_can71_2020$extrapolation),
                          na.color = "transparent")

  # Leaflet map
   training_data_map <- leaflet() %>%
    
  # Setting view Canada
   setView(lng = -82, lat = 50, zoom = 4)%>%
     
  # Adding background map
    addProviderTiles(
    "Esri.WorldImagery",'https://server.arcgisonline.com/ArcGIS/rest/services/World_Imagery/MapServer/tile/{z}/{y}/{x}')%>%
 
  # Adding BCR map
    addPolygons(data=bcr_boundary_wgs84,
                weight=1,
                fillColor = '0',
                color = "white",
                label=bcr_boundary_wgs84$subUnit,
                 group= "Regions"
                )%>%
  # Adding species data
    addCircleMarkers(data=terra::project(locations_can71, "epsg:4326"),
                     fillColor = ~pal(locations_can71$count),
                     weight = 0.5,
                     radius = 2, 
                     color = ~pal(locations_can71$count),
                     group="Observations")%>%
     
  # Adding legend species data
    addLegend('bottomright', pal = pal, values = locations_can71$count,
            title = "Count", opacity = 1, group="Species_data")%>%
     
  # Adding raster PREDICTIONS
    addRasterImage(terra::project(BBWA_can71_2020$mean, "epsg:4326"),
                   colors=pal2,opacity=0.8,
                   group="Predictions")%>%
    
     addLegend(pal = pal2,
               values = values(terra::project(BBWA_can71_2020$mean, "epsg:4326")),
               title = "Predictions (Mean-quantiles)",
               position = "bottomright",
               group = "Predictions")%>%
     

  # Adding raster UNCERTAINTY
    addRasterImage(terra::project(BBWA_can71_2020$cv, "epsg:4326"),
                   colors=pal3, opacity=0.8, group="Uncertainty")%>%
    
    addLegend(pal = pal3, values = values(BBWA_can71_2020$cv),
               title = "Uncertainty(CV-Quantiles)",
               position = "bottomright",
               layerId = "Uncertainty",
              group = "Uncertainty")%>%
     
 # Adding raster EXTRAPOLATION
     addRasterImage(terra::project(BBWA_can71_2020$extrapolation, "epsg:4326"),
                    colors=pal4,opacity=0.8, group="Extrapolation")%>%
     
     addLegend(pal = pal4, values = values(BBWA_can71_2020$extrapolation),
                title = "Extrapolation",
                position = "bottomright",
                layerId = "Extrapolation",
               group = "Extrapolation")%>%
  

   # add measurement tool
    addMeasure(primaryLengthUnit = "kilometers",
               secondaryLengthUnit = 'miles',
               primaryAreaUnit = "hectares",
               secondaryAreaUnit="acres",
               position = 'topleft')%>%
  
 # Adding layer control
    addLayersControl(options = layersControlOptions(collapsed = T),
                       position ="topleft",
                       overlayGroups = c( "Regions",
                                          "Species_data",
                                          "Predictions",
                                          "Uncertainty",
                                          "Extrapolation"))%>%
                       hideGroup(c("Predictions", "Uncertainty", "Extrapolation"))
                     


```

```{r}
  training_data_map
```
